Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2021-09-29 methods_paper/fabl-paper.tex
update.

The hashing procedure described above considerably reduces the memory
needed to store the comparison information: instead of storing
\(n_A \times n_B\) comparison vectors, which are relatively long under
either the Fellegi Sunter or our modified framework, we only store the
\(P\) unique vectors, and then \(n_A \times n_B\) scalar quantities
relating record pairs to those vectors. However, even storing these
\(n_A \times n_B\) scalar labels can become burdensome with large data.
Worse, the overwhelming majority of these labels relate to record pairs
that are clear non-matches.

To address this, we propose a new strategy called \emph{storage	efficient indexing} (SEI). In standard indexing, the modeler decides a
certain criteria that they expect all true matching pairs to satisfy,
and a priori label any record pairs that do not meet that criteria as
non-matching. For example, one might only consider pairs with a certain
similarity score on a field deemed to be important (like first name), or
only pairs with exact matching on a specified number of fields. While
generally chosen to be be quite loose, establishing these criteria
requires knowledge of the problem and invites room for human error. We
propose a method of reducing the comparison space (and reducing storage
requirements) without these drawbacks. Note that all records of the same
agreement pattern have the same probability when sampling \(Z_j\).
Therefore we know that records belonging to an \(h_p\) such that
\(H_{p_j}\) is large are very unlikely to be sampled consistently enough
to be deemed a match through the Bayes estimate, even without
considering the form of the agreement pattern itself.

In SEI, rather than store all of these unlikely record labels, we choose
to store only a small number \(R\) of them. Posterior calculations still
attribute the appropriate weight to all records through the summary
statistics \(H_p\), and \(H_{p_j}\). Rather than storing
\(n_A \times n_B\) record labels, SEI allows us to store at most
\(n_B \times P \times R\) labels, regardless of how large \(n_A\) is.

Lastly, for large data, we can partition the two datasets \(A\) and
\(B\) into smaller blocks \(\{A_m\}\) and \(\{B_m\}\) for more
manageable computations. On a single machine, we can read-in data
sequentially, conduct hashing, compress information through SEI, and
delete the original data from memory before continuing with the next
chunk of data. With multiple cores or multiple machines, this can be
done in parallel. Thus the combination of hashing, SEI, and partitioning
allows us to conduct linkage tasks over much larger datasets.

>>>>

The hashing procedure described above considerably reduces the memory
needed to store the comparison information. That is, instead of storing
\(n_1 \times n_1\) comparison vectors, which are relatively long under
either the Fellegi Sunter or our modified framework, we only store the
\(P\) unique vectors, and then \(n_1 \times n_2\) scalar quantities
relating record pairs to those vectors. However, even storing these
\(n_1 \times n_2\) scalar labels can become burdensome with large data.
Worse, the overwhelming majority of these labels relate to record pairs
that are clear non-matches.

To address this issue, we propose a new method called \emph{storage	efficient indexing} (SEI). In standard indexing, one decides a certain criteria that they expect all true matching pairs to satisfy. In addition, one decides 
a priori to label any record pairs that do not meet that criteria as
non-matches. For example, one might only consider pairs with a certain
similarity score on a field deemed to be important (like first name), or
only pairs with exact matching on a specified number of fields. While
generally chosen to be be quite loose, establishing these criteria
requires knowledge of the problem and invites room for human error. 
To improve upon this, we propose to reduce the comparison space and reduce the storage requirements while avoiding the drawbacks of standard indexing.

Observe that all records of the same
agreement pattern have the same probability when sampling \(Z_j\).
Therefore we know that records belonging to an \(h_p\) such that
\(H_{p_j}\) is large are very unlikely to be sampled consistently enough
to be deemed a match through the Bayes estimate, even without
considering the form of the agreement pattern itself.

In SEI, rather than store all of these unlikely record labels, we choose
to store only a small number \(R\) of them. Posterior calculations still
attribute the appropriate weight to all records through the summary
statistics \(H_p\), and \(H_{p_j}\). Rather than storing
\(n_1 \times n_2\) record labels, SEI allows us to store at most
\(n_2 \times P \times R\) labels, regardless of how large \(n_1\) might be.

Lastly, for large data, we can partition the two datasets \(\bm{X}_1\) and
\(\bm{X}_2\) into smaller blocks \(\{\bm{X}_{1m}\}\) and \(\{\bm{X}_{2m}\}\) for more
manageable computations. On a single machine, we can read-in data
sequentially, conduct hashing, compress information through SEI, and
delete the original data from memory before continuing with the next
chunk of data. With multiple cores or multiple machines, this can be
done in parallel. Thus, the combination of hashing, SEI, and partitioning
allows us to conduct linkage tasks over much larger datasets.

\textcolor{red}{Brian: how large can we go and how fast can we go?}
